P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2023-12-20 manuscript.tex
Add references and fix few parts in the text
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

P4IRS can generate several kernel flavours of the same potential such as Linked Cells and Verlet Lists, where data layouts for properties and specific arrays (such as the neighbor-lists) can be switched to the case with best performance.
A runtime interface in C++ is also available and can be used to integrate P4IRS with other simulation tools, which is particularly important for us to be able to integrate P4IRS simulations with our multi-physics framework waLBerLa.
%as an example we show how P4IRS can use the load-balancing communication algorithm from waLBerLa to compute simulations with a more heterogeneous distribution of particles.

@@

The goals of this paper are summarized as follows:
\begin{itemize}
	\item We present our code generator tool P4IRS, describing how it works and how it can be used to generate efficient particle simulation codes.
	\item We discuss the advantages of using P4IRS to achieve performance-portability, also comparing it with other previous approaches such as tinyMD and MESA-PD.
	\item We describe P4IRS interfaces to allow its generated code to be integrated with other simulation tools, especifically in the context of domain-partitioning and communication.
	\item We provide performance and scaling results for Intel Ice Lake, AMD Milan and NVIDIA A40 and A100 hardwares, to demonstrate the state-of-the-art performance and scaling capabilities of the applications generated by P4IRS.
\end{itemize}

@@

In MD, there exist many packages with distinct strategies and simulation fields.
LAMMPS is a molecular package for the material-modeling field which uses the standard Verlet Lists algorithm to optimize the non-bonded short-range force calculations.
It offers many interatomic potentials, each one with different kernel implementations for a specific target (like CPU, GPU, Kokkos).
GROMACS is another package for MD with focus on bio-sciences simulation which uses its own Cluster Pair algorithm to optimize the short-range forces computation.
It does not support as many potentials as LAMMPS, but its algorithm tends to be significantly faster than the Verlet Lists since it employs SIMD parallelism more efficiently.
Furthermore, the optimal GROMACS kernel implementations rely on SIMD intrinsics to attain higher performance, which makes them harder to be maintained and ported to other architectures.
Also, the GPU algorithm in GROMACS differs from the CPU, since it relies on a strategy named ``super-clustering'' to reduce the amount of redundant computations caused by adjusting the cluster sizes to fit the SIMD width in the GPU.
Other MD packages such as HOOMD-blue, NAMD, AMBER and CHARMM also contain dedicated and hand-tuned implementations in general purpose languages for each supported backend.
Each version is tested, maintained, debugged and optimized independently.

Packages that support DEM simulations are in general more flexible (since simulation with more particle properties such as rotations and shapes are needed), however they usually do not have the same optimization techniques as the MD simulation packages.
The reason is most likely that it is not so feasible to support flexibility with sophisticated optimization algorithms as the complexity of the code grows considerably.
Examples of such packages are the LIGGGHTS DEM (based on LAMMPS), GranOO and YADE.
To mitigate such limitations, a few code generation approaches were developed for both MD and DEM.

@@

OpenMM is a \ac{MD} package that makes use of Python as a front-end for simulation protocols and file I/O, it contains a C++ API for forces and integrators as well as optimized kernel implementations for CUDA, C++ and OpenCL.
Integrators and forces can be implemented by using mathematical descriptions in OpenMM, but its use cases are restricted to \ac{MD} simulations and several features required for a DEM simulation or other types of particles simulation (such as contact history and different shapes) are not available.

MESA-PD is a general particle dynamics framework that is built-in the waLBerla framework for the simulation of particles and fluids.
Its design principle of separation of code and data allows to introduce arbitrary interaction models with ease, which makes it also useful for \ac{MD}.
By using its code generation approach using Jinja templates it can be adapted very effectively to any simulation scenario.
As successor of the pe rigid particle dynamics framework it inherits its scalability and advanced load balancing functionalities.
One of the limitations of MESA-PD is that it does not support simulations in the GPU, and since its code generation approach is based on Jinja templates, the effort for implementing new functionalities may not be lower compared to other simulation packages without code generation techniques (for instance, most of its force kernels are not generated but rather implemented in C++).

In previous work we developed tinyMD, a proxy-app (also based on miniMD) created to evaluate the portability of MD and DEM applications with the AnyDSL framework.
TinyMD uses higher-order functions to abstract device iteration loops, data layouts and communication strategies, providing a domain-specific library to implement pair-wise interaction kernels that execute efficiently on multi-CPU and multi-GPU targets.
One of tinyMD's limitation is that its flexibility and optimization settings are restricted to what can be achieve through higher-order functions and the AnyDSL compilation framework (the generated LLVM IR restricts the backend compilation to the Clang compiler), overcoming such limitation is one of the reasons we develop P4IRS.

@@

Several performance investigation researches for MD systems exist, most of them using MD packages as a black-box and adjusting different simulation parameters to analyse their implications in the overall performance.
In some cases, however, in-depth performance studies were made to gain a broader understanding on efficient ways to achieve SIMD vectorization with SSE/AVX instructions.
Most in-depth studies are performed using mini-apps or proxy-apps, with miniMD being worth to mention as it provides a simple copper face-cubic-centered lattice simulation using the kernels extracted from LAMMPS.
Our performance-focused prototyping harness MD-Bench is a toolbox developed for performance-engineering \ac{MD} non-bonded short-range force calculation algorithms, with focus on in-core execution on CPUs and GPUs.
It currently supports the Verlet Lists and Cluster Pair algorithms from LAMMPS and GROMACS, respectively.
One of our goals is to use the knowledge obtained with MD-Bench and integrate it into P4IRS to achieve state-of-the-art performance and flexibilty altogether in particle simulations.

@@

Where $k^{n}$ is the spring stiffness, $\xi^{n}$ the normal velocity damping and $\mu$ the friction.
The update rule \autoref{eq:update_rule} and modifications in \autoref{eq:linear_spring_dashpot_tangential} were properly included to fix frictional contact accordingly.
The update rule is carried out every time-step the contact persists, and therefore a contact history data structure is needed to properly compute it from the algorithmic point of view.

@@

\begin{figure*}[htb]
    \centering
    \subfigure[Linked Cells]{\label{fig:linked_cells}\includegraphics[width=0.49\textwidth]{linked_cells.pdf}}
    \subfigure[Verlet Lists]{\label{fig:verlet_lists}\includegraphics[width=0.49\textwidth]{verlet_lists.pdf}}
    \caption{Comparison between (a) Linked Cells and (b) Verlet Lists algorithms, in each case, the blue particles are used as the candidates to compute the forces with the red particle. In the Linked Cells, all particles in the neighbor cells are considered (considering the size of the cell is higher or equal than the cutoff radius). In the Verlet Lists, a list of atoms is created by evaluating which particle distances are smaller than the cutoff radius plus a "skin" value, which represents a sphere in the 3D case.}
\end{figure*}

@@

The Linked Cells results in computing only forces for particles within a region that is discretized by their current and neighbor cells/bins, which can still results in a significant amount of redundant computations.
In order avoid even more redundant computations, the Verlet Lists strategy can be used jointly with Linked Cells.
The Verlet Lists (see Figure~\ref{fig:verlet_lists}) constructs a list of neighbor particles for every particle, and this list can be created by using the Linked Cells structure, and doing particle-particle distance comparison to evaluate wether a particle should be added into the list.
At the end, the region for each particle search is defined by a sphere.
Note that for the neighbor-lists construction to be worthwhile, it should be used more than once force calculation, hence neighbor-lists are not created at every simulation time-step, and a "skin" can be added to the cutoff radius to extend the neighbor-search and avoid missing computations from atoms that end up entering the cut-off region.

Despite avoiding computations from long-distance particles, another common optimization is to take advantage of Newton's Third Law and computing forces in only one direction.
This means that only half of particle-pairs are traversed during force computation, and forces are decreased for the neighbor atom during their computation.
Ideally, this should lead to a factor of two speedup, but this strategy has a few drawbacks: (a) parallelism is harmed since decreasing force of neighbor atoms (j) requires atomic operations to avoid race conditions and (b) gather and scatter instructions are needed to build up the SIMD registers with forces of the neighbor atoms (j), which results in scattered accesses and extra instructions.
At the end, this strategy is likely to be beneficial for more arithmetic-intensive kernels, and the trade-off between parallelism and less arithmetic operations must be assessed.

@@

\begin{itemize}
  \item \textbf{Fritz:} 992 compute nodes, each one with two Intel Xeon Platinum 8360Y “Ice Lake” processors (36 cores per chip) running at a base frequency of 2.4 GHz and 54 MB Shared L3 cache per chip, 256 GB of DDR4 RAM.
  \item \textbf{Alex:} 44 GPGPU nodes, each with two AMD EPYC 7713 “Milan” processors (64 cores per chip) running at 2.0 GHz with 256 MB Shared L3Cache per chip, 512 GB of DDR4-RAM, eight Nvidia A40 (each with 48 GB DDR6 @ 696 GB/s; 37.42 TFlop/s in FP32), 25 GbE, and 7 TB on local NVMe SSDs.
\end{itemize}

We perform single-core CPU experiments in both clusters, and refer to both CPUs listed above as Ice Lake and Milan, respectively.
For single GPU experiments, we use the referred A40 Nvidia GPUs, as well as the Nvidia A100 also available in the Alex cluster.

Experiments for \ac{MD} use the standard Copper FCC lattice testcase from miniMD/MD-Bench, and we use our results from MD-Bench as a performance reference since it is a well optimized code for \ac{MD}.
The \ac{DEM} test case is a "Settling Spheres" benchmark, where several spheres fall into a "ground" (half-space), thus generating a "bed" of spherical particles.
Due to its multiple shapes and dynamics that consists of many interactions among spheres and half-spaces, it should provide a good reference for DEM simulations.
We compare our single-core results from P4IRS with MESA-PD using 2 MPI ranks since it is not possible to run this setup in MESA-PD using a single rank.
GPU results are not available for MESA-PD as it is not capable of generating GPU code, but we show the performance benefits when using GPUs for P4IRS generated codes.

Scaling results are also shown for demostrating P4IRS scaling capabilities.
For this we use regular domain-partitioning since these experiments do not require sophisticated load-balancing algorithms (Copper lattice is an homogeneous testcase in all dimensions, and Settling Spheres can obtain good weak-scaling by regularly partitioning the domain in X and Y dimensions).
We use double floating-point precision and display results when enabling and disabling Newton's 3rd Law optimization (half neighbor-lists).
For \ac{MD} we show results for the Verlet Lists case since it is widely employed in such simulations, and for \ac{DEM} we stick to Linked Cells as this is the algorithm used in MESA-PD.

\begin{figure*}[tb]
    \centering
    \subfigure[t = 0]{\label{fig:md_ts0}\includegraphics[width=0.32\textwidth]{md_0.png}}
    \subfigure[t = 100]{\label{fig:md_ts100}\includegraphics[width=0.32\textwidth]{md_100.png}}
    \subfigure[t = 200]{\label{fig:md_ts200}\includegraphics[width=0.32\textwidth]{md_200.png}}
    \caption{The MD simulation example used in this paper, which is a Copper FCC lattice simulation with homogeneous distribution. Forces are computed with the van der Waals potential and PBC is used in all directions. In average, every atom contains roughly 76 neighbors through the whole simulation (200 time-steps).}
    \label{fig:settling_spheres}
\end{figure*}

@@

\autoref{fig:md_single_core_runtimes} display the runtime measurements for P4IRS and MD-Bench simulations of the Copper FCC lattice case with $32^{3}$ unit cells (4 atoms per unit-cell).
It is noticeable that the kernels generated by P4IRS achieve competitive results to MD-Bench, a hand-tuned implementation for MD cases, which demonstrates P4IRS capabitilies to generate fast code.
In Ice Lake, The performance is slightly better for MD-Bench using \ac{FN}, but P4IRS provide a slightly faster result when using \ac{HN}.
P4IRS code tend to be build neighbor-lists faster but the force computation is slower, but runtime differences vary in the range of 10\% for the measured runtime, which is still not too significant and can vary when using different compilers and flags.
In Milan, P4IRS is faster than MD-Bench for the \ac{FN}case, and results are pretty similar when using \ac{HN} case.
The force calculation is still slower for P4IRS with \ac{HN} when compared to MD-Bench, but this is compensated by the less time spent building the neighbor lists.

@@

\autoref{fig:md_weak_scaling_alex} shows the results for P4IRS weak-scaling in the Alex supercomputer for both nodes with A40 and A100 GPUs.
Each node simulates a system of $100^3$ unit cells, with roughly 500.000 atoms per GPU.
Note there is a significant overhead for communication when using GPUs (compared to solving the problem in a single GPU), which can be attributed to the time do transfer data from the GPUs through the communication channels.
The results for scaling up to 8 nodes (64 GPUs) is good but not perfect (93\% efficiency for \ac{HN} in A100, and 83-85\% efficiency across other cases), one reason for this is that the Alex cluster has significant impact with its interconnect interface among different nodes, which significantly increases the time for the communication.
Running simulations with higher workload on GPUs for the force computation or using different GPU clusters with better interconnection interfaces should improve the weak-scaling results.

@@

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-num-names} 
%%  \bibliography{<your bibdatabase>}

@@

\begin{thebibliography}{00}

@@

\bibitem[ ()]{}

\end{thebibliography}
\end{document}

>>>>

P4IRS can generate several kernel flavours of the same potential such as Linked Cells \cite{linkedcells} and Verlet Lists \cite{verletlists}, where data layouts for properties and specific arrays (such as the neighbor-lists) can be switched to the case with best performance.
A runtime interface in C++ is also available and can be used to integrate P4IRS with other simulation tools, which is particularly important for us to be able to integrate P4IRS simulations with our multi-physics framework waLBerLa \cite{walberla1,walberla2}.
%as an example we show how P4IRS can use the load-balancing communication algorithm from waLBerLa to compute simulations with a more heterogeneous distribution of particles.

@@

The goals of this paper are summarized as follows:
\begin{itemize}
	\item We present our code generator tool P4IRS, describing how it works and how it can be used to generate efficient particle simulation codes.
	\item We discuss the advantages of using P4IRS to achieve performance-portability, also comparing it with other previous approaches such as tinyMD \cite{tinymd} and MESA-PD \cite{mesapd1,mesapd2}.
	\item We describe P4IRS interfaces to allow its generated code to be integrated with other simulation tools, especifically in the context of domain-partitioning and communication.
	\item We provide performance and scaling results for Intel Ice Lake, AMD Milan and NVIDIA A40 and A100 hardwares, to demonstrate the state-of-the-art performance and scaling capabilities of the applications generated by P4IRS.
\end{itemize}

@@

In MD, there exist many packages with distinct strategies and simulation fields.
LAMMPS \cite{lammps1,lammps2} is a molecular package for the material-modeling field which uses the standard Verlet Lists algorithm to optimize the non-bonded short-range force calculations.
It offers many interatomic potentials, each one with different kernel implementations for a specific target (like CPU, GPU or Kokkos \cite{kokkos}).
GROMACS \cite{gromacs1,gromacs2} is another package for MD with focus on bio-sciences simulation which uses its own Cluster Pair algorithm to optimize the short-range forces computation.
It does not support as many potentials as LAMMPS, but its algorithm tends to be significantly faster than the Verlet Lists since it employs SIMD parallelism more efficiently.
Furthermore, the optimal GROMACS kernel implementations rely on SIMD intrinsics to attain higher performance, which makes them harder to be maintained and ported to other architectures.
Also, the GPU algorithm in GROMACS differs from the CPU, since it relies on a strategy named ``super-clustering'' to reduce the amount of redundant computations caused by adjusting the cluster sizes to fit the SIMD width in the GPU.
Other MD packages such as HOOMD-blue \cite{hoomdblue}, NAMD \cite{namd}, AMBER \cite{amber1,amber2} and CHARMM \cite{charmm} also contain dedicated and hand-tuned implementations in general purpose languages for each supported backend.
Each version is tested, maintained, debugged and optimized independently.

Packages that support DEM simulations are in general more flexible (since simulation with more particle properties such as rotations and shapes are needed), however they usually do not have the same optimization techniques as the MD simulation packages.
The reason is most likely that it is not so feasible to support flexibility with sophisticated optimization algorithms as the complexity of the code grows considerably.
Examples of such packages are the LIGGGHTS DEM \cite{liggghts} (based on LAMMPS), MercuryDPM \cite{mercurydpm} and YADE \cite{yade}.
To mitigate such limitations, a few code generation approaches were developed for both MD and DEM.

@@

OpenMM \cite{openmm} is a \ac{MD} package that makes use of Python as a front-end for simulation protocols and file I/O, it contains a C++ API for forces and integrators as well as optimized kernel implementations for CUDA, C++ and OpenCL.
Integrators and forces can be implemented by using mathematical descriptions in OpenMM, but its use cases are restricted to \ac{MD} simulations and several features required for a DEM simulation or other types of particles simulation (such as contact history and different shapes) are not available.

MESA-PD \cite{mesapd3} is a general particle dynamics framework that is built-in the waLBerla framework for the simulation of particles and fluids.
Its design principle of separation of code and data allows to introduce arbitrary interaction models with ease, which makes it also useful for \ac{MD}.
By using its code generation approach using Jinja templates it can be adapted very effectively to any simulation scenario.
As successor of the pe rigid particle dynamics framework it inherits its scalability and advanced load balancing functionalities.
One of the limitations of MESA-PD is that it does not support simulations in the GPU, and since its code generation approach is based on Jinja templates, the effort for implementing new functionalities may not be lower compared to other simulation packages without code generation techniques (for instance, most of its force kernels are not generated but rather implemented in C++).

In previous work we developed tinyMD, a proxy-app (also based on miniMD) created to evaluate the portability of MD and DEM applications with the AnyDSL \cite{anydsl1,anydsl2} framework.
TinyMD uses higher-order functions to abstract device iteration loops, data layouts and communication strategies, providing a domain-specific library to implement pair-wise interaction kernels that execute efficiently on multi-CPU and multi-GPU targets.
One of tinyMD's limitation is that its flexibility and optimization settings are restricted to what can be achieve through higher-order functions and the AnyDSL compilation framework (the generated LLVM IR restricts the backend compilation to the Clang compiler), overcoming such limitation is one of the reasons we develop P4IRS.

@@

Several performance investigation researches for MD systems exist, most of them using MD packages as a black-box and adjusting different simulation parameters to analyse their implications in the overall performance.
In some cases, however, in-depth performance studies were made to gain a broader understanding on efficient ways to achieve SIMD vectorization with SSE/AVX instructions \cite{mdsimd}.
Most in-depth studies are performed using mini-apps or proxy-apps, with miniMD being worth to mention as it provides a simple copper face-cubic-centered lattice simulation using the kernels extracted from LAMMPS.
Our performance-focused prototyping harness MD-Bench \cite{mdbench1,mdbench2} is a toolbox developed for performance-engineering \ac{MD} non-bonded short-range force calculation algorithms, with focus on in-core execution on CPUs and GPUs.
It currently supports the Verlet Lists and Cluster Pair algorithms from LAMMPS and GROMACS, respectively.
One of our goals is to use the knowledge obtained with MD-Bench and integrate it into P4IRS to achieve state-of-the-art performance and flexibilty altogether in particle simulations.

@@

Where $k^{n}$ is the spring stiffness, $\xi^{n}$ the normal velocity damping and $\mu$ the friction.
The update rule \autoref{eq:update_rule} and modifications in \autoref{eq:linear_spring_dashpot_tangential} were properly included to fix frictional contact accordingly \cite{dem1,dem2}.
The update rule is carried out every time-step the contact persists, which from the algorithmic point of view means that a contact history data structure is needed to properly compute these interactions.

@@

\begin{figure*}[htb]
    \centering
    \subfigure[Linked Cells]{\label{fig:linked_cells}\includegraphics[width=0.40\textwidth]{linked_cells.pdf}}
    \subfigure[Verlet Lists]{\label{fig:verlet_lists}\includegraphics[width=0.40\textwidth]{verlet_lists.pdf}}
    \caption{Comparison between (a) Linked Cells and (b) Verlet Lists algorithms, in each case, the blue particles are used as the candidates to compute the forces with the red particle. In the Linked Cells, all particles in the neighbor cells are considered (considering the size of the cell is higher or equal than the cutoff radius). In the Verlet Lists, a list of atoms is created by evaluating which particle distances are smaller than the cutoff radius plus a "skin" value, which represents a sphere in the 3D case.}
\end{figure*}

@@

The Linked Cells results in computing only forces for particles within a region that is discretized by their current and neighbor cells/bins, which can still results in a significant amount of redundant computations.
In order to avoid even more redundant computations, the Verlet Lists strategy can be used jointly with Linked Cells.
The Verlet Lists (see Figure~\ref{fig:verlet_lists}) constructs a list of neighbor particles for every particle, and this list can be created by using the Linked Cells structure, evaluating the distances among particles within close bins and then using the distance to determine whether such particles should be added into the list.
In the Verlet Lists case, the region for searching for neighbor particles is defined by a sphere, where the center of the sphere is the position of the particle and the radius is the cutoff radius plus a "skin".
The "skin" is used to avoid building the neighbor-lists at every time-step, it extends the neighbor-search and avoid missing computations from atoms that end up entering the cut-off region in subsequent time-steps in which the neighbor-lists is not reconstructed.

Despite avoiding computations from long-distance particles, another common optimization is to take advantage of Newton's Third Law and computing forces in only one direction.
This means that only half of particle-pairs are traversed during force computation, and forces are decreased for the neighbor atom during their computation.
Ideally, this should lead to a factor of two speedup, but this strategy has a few drawbacks: (a) parallelism is harmed since decreasing force of neighbor atoms requires atomic operations to avoid race conditions and (b) when the code is vectorized with AVX/SSE intrinsics, gather and scatter instructions are needed to build up the SIMD registers with forces of the neighbor atoms, which results in scattered memory accesses and extra instructions.
At the end, this strategy is likely to be beneficial for more arithmetic-intensive kernels, and the trade-off between parallelism and less arithmetic operations must be assessed.

@@

\begin{itemize}
  \item \textbf{Fritz:} Each compute node has two Intel Xeon Platinum 8360Y “Ice Lake” processors (36 cores per chip) running at a base frequency of 2.4 GHz and 54 MB Shared L3 cache per chip, 256 GB of DDR4 RAM.
  \item \textbf{Alex:} Each compute node has two AMD EPYC 7713 “Milan” processors (64 cores per chip) running at 2.0 GHz with 256 MB Shared L3Cache per chip, 512 GB of DDR4-RAM, eight GPUs are available per node which can be either Nvidia A40 (with 48 GB DDR6 @ 696 GB/s; 37.42 TFlop/s in FP32) or A100 (each 80 GB HBM2 @ 2,039 GB/s; HGX board with NVLink; 9.7 TFlop/s in FP64 or 19.5 TFlop/s in FP32)
\end{itemize}

We perform single-core CPU experiments in both clusters, and refer to both CPUs listed above as Ice Lake and Milan, respectively.
For single GPU experiments, we use both A40 and A100 Nvidia GPUs.
The icc (ICC) 2021.10.0 20230609 compiler was used for generating the CPU code, with the \emph{-Ofast -xCORE-AVX512 -qopt-zmm-usage=high} flags for Ice Lake (AVX512) and \emph{-Ofast -march=core-avx2} flags for Milan (AVX2).
For CUDA, we use the NVCC compiler v12.1.r12.1/compiler.32688072, with same flags as the ones used for the Milan system.
We make use of the Likwid toolkit V5.3 \cite{likwid} to measure HPM counters in the target CPU (Ice Lake) for the single core experiments, where P4IRS is able to properly generate the instrumentation markers and MD-Bench kernels are already instrumented.

Experiments for \ac{MD} use the standard Copper FCC lattice (see \autoref{fig:copper_fcc_lattice}) testcase from miniMD/MD-Bench, and we use our results from MD-Bench as a performance reference since it is a well optimized code for \ac{MD}.
The \ac{DEM} test case is a "Settling Spheres" benchmark (see \autoref{fig:settling_spheres}), on which several spheres fall into a "ground" (half-space), thus generating a "bed" of spherical particles.
Due to its multiple shapes and dynamics that consists of many interactions among spheres and half-spaces, it should provide a good reference for DEM simulations.
We compare our single-core results from P4IRS with MESA-PD using 2 MPI ranks since it is not possible to run this setup in MESA-PD using a single rank.
GPU results are not available for MESA-PD as it is not capable of generating GPU code, but we show GPU performance results for P4IRS.

We also show weak-scaling results to demostrate P4IRS scaling capabilities.
For this we use regular domain-partitioning since these experiments do not require sophisticated load-balancing algorithms (Copper lattice is an homogeneous testcase in all dimensions, and Settling Spheres can obtain good weak-scaling by regularly partitioning the domain in X and Y dimensions).
We use double floating-point precision and display results when enabling and disabling Newton's 3rd Law optimization (half neighbor-lists).
For \ac{MD} we show results for the Verlet Lists case since it is widely employed in such simulations, and for \ac{DEM} we stick to Linked Cells as this is the algorithm used in MESA-PD.

\begin{figure*}[tb]
    \centering
    \subfigure[t = 0]{\label{fig:md_ts0}\includegraphics[width=0.32\textwidth]{md_0.png}}
    \subfigure[t = 100]{\label{fig:md_ts100}\includegraphics[width=0.32\textwidth]{md_100.png}}
    \subfigure[t = 200]{\label{fig:md_ts200}\includegraphics[width=0.32\textwidth]{md_200.png}}
    \caption{The MD simulation example used in this paper, which is a Copper FCC lattice simulation with homogeneous distribution. Forces are computed with the van der Waals potential and PBC is used in all directions. In average, every atom contains roughly 76 neighbors through the whole simulation (200 time-steps).}
    \label{fig:copper_fcc_lattice}
\end{figure*}

@@

\autoref{fig:md_single_core_runtimes} display the runtime measurements for P4IRS and MD-Bench simulations of the Copper FCC lattice case with $32^{3}$ unit cells (4 atoms per unit-cell).
It is noticeable that the kernels generated by P4IRS achieve competitive results to MD-Bench, a hand-tuned implementation for MD cases, which demonstrates P4IRS capabilities to generate fast code.
For Ice Lake, performance is slightly better for MD-Bench using \ac{FN}, but P4IRS provide a slightly faster result when using \ac{HN}.
P4IRS code tends to be build neighbor-lists faster but the force computation is slower, but runtime differences vary in the range of 10\% for the measured runtime, which is still not too significant and can vary when using different compilers and flags.
In Milan, P4IRS is faster than MD-Bench for the \ac{FN} case, and results are pretty similar when using \ac{HN} case.
The force calculation is still slower for P4IRS with \ac{HN} when compared to MD-Bench, but this is compensated by the less time spent building the neighbor lists.

@@

\autoref{fig:md_weak_scaling_alex} shows the results for P4IRS weak-scaling in the Alex supercomputer for both nodes with A40 and A100 GPUs.
Each node simulates a system of $100^3$ unit cells, with roughly 500.000 atoms per GPU.
Note there is a significant overhead for communication when using GPUs (compared to solving the problem in a single GPU), which can be attributed to the time spent transferrng data from the GPUs through the communication channels and then back to the other GPU.
The results for scaling up to 8 nodes (64 GPUs) are good but not perfect (93\% efficiency for \ac{HN} in A100, and 83-85\% efficiency across other cases), one reason for this is that the Alex cluster has significant impact with its interconnect interface among different nodes, which significantly increases the time for the communication.
Running simulations with higher workload on GPUs for the force computation or using different GPU clusters with better interconnection interfaces should improve the weak-scaling results.
In any case, however, it is noticeable that P4IRS can generate code that scales up within GPU clusters.

@@

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%

\bibliographystyle{elsarticle-num-names} 
\bibliography{literature}

@@

%\begin{thebibliography}{00}

@@

%\bibitem[ ()]{}

%\end{thebibliography}
\end{document}
